---
title: "API Reference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{API Reference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette provides a complete reference for the RIBITSr package functions,
RIBITS API endpoints, and data fields.

## Function Reference

### Unified API Functions

These are the recommended functions for most use cases.

#### `rb_get()` - Main Data Retrieval

```{r rb-get}
rb_get(
  type = NULL,           # Data type: "banks", "ilf", "umbrellas", "wqt"
  id = NULL,             # Specific ID(s) for detailed data
  state = NULL,          # State filter: "CA", "OR", "TX"
  district = NULL,       # USACE district: "Portland", "Sacramento"
  field_office = NULL,   # FWS field office
  noaa_region = NULL,    # NOAA region
  kind = NULL,           # Bank type: "Standard", "ILF", "Umbrella", "NRDA"
 status = NULL,         # Status: "Approved", "Pending", "Terminated"
  ledger = FALSE,        # Include transaction history
  footprint = FALSE,     # Include footprint geometry
  service_area = FALSE,  # Include service area geometry
  contacts = FALSE       # Include contact information
)

# Examples
rb_get()                                    # Show help
rb_get("banks", state = "OR")               # List Oregon banks
rb_get("banks", id = 17, ledger = TRUE)     # Get bank with transactions
rb_get("ilf", district = "Portland")        # List Portland ILF programs
```

#### `rb_epa()` - EPA ArcGIS Spatial Data

```{r rb-epa}
rb_epa(
  layer = NULL,          # Layer: "banks", "footprints", "service_areas", etc.
  state = NULL,          # State filter
  district = NULL,       # USACE district filter
  bank_ids = NULL,       # Vector of bank IDs
  program_ids = NULL,    # Vector of program IDs (for ILF layers)
  status = NULL,         # Status filter
  kind = NULL,           # Bank type filter
  where = NULL           # Custom SQL WHERE clause
)

# Examples
rb_epa()                                     # Show available layers
rb_epa("banks", state = "OR")                # Oregon banks
rb_epa("footprints", bank_ids = c(17, 100))  # Specific footprints
rb_epa("ilf_service_areas")                  # All ILF service areas
```

#### `rb_download_report()` - Direct CSV Downloads

Download RIBITS reports programmatically (no browser needed):

```{r rb-download}
rb_download_report(
  report_type,           # Report type (see rb_report_types())
  download_dir = "data/ribits_reports",
  filename = NULL,       # Auto-generated if NULL
  reset_filters = TRUE   # Reset report filters to get all data
)

# See available reports
rb_report_types()

# Examples
rb_download_report("banks_sites", download_dir = "data/")
rb_download_report("credit_classification", download_dir = "data/")
rb_download_report("ledger_transactions", download_dir = "data/")
```

#### `rb_read()` - Read Downloaded CSVs

```{r rb-read}
rb_read(
  path,                  # Path to CSV file
  type = NULL            # Optional: force specific parser (usually auto-detected)
)

# Examples
data <- rb_read("data/banks_sites.csv")
credits <- rb_read("data/potential_credits.csv")  # Auto-detects special parser
```

#### `rb_near()` - Location-Based Search

```{r rb-near}
rb_near(
  lat,                   # Latitude (decimal degrees)
  lon,                   # Longitude (decimal degrees)
  type = "banks"         # "banks" or "programs"
)

# Example
nearby <- rb_near(lat = 45.5152, lon = -122.6784)
```

### Extraction Functions

#### `rb_extract()` - Extract Nested Data

```{r rb-extract}
rb_extract(
  data,                  # Bank/program object from rb_get()
  what                   # What to extract (see below)
)

# Available extractions:
# - "ledger" - Transaction history
# - "contacts" - All contacts
# - "sponsors" - Bank sponsors
# - "pocs" - Points of contact
# - "managers" - Bank managers
# - "footprint" - Footprint geometry
# - "service_area" - Service area geometry

# Example
bank <- rb_get("banks", id = 17, ledger = TRUE)
ledger <- rb_extract(bank, "ledger")
```

#### `rb_bulk_ledger()` - Bulk Ledger Extraction

```{r rb-bulk}
rb_bulk_ledger(
  bank_ids,              # Vector of bank IDs
  progress = TRUE        # Show progress bar
)

# Example
ledgers <- rb_bulk_ledger(bank_ids = 1:100)
```

### Legacy Functions

These functions still work but the unified API is recommended.

| Legacy Function | Unified Equivalent |
|-----------------|-------------------|
| `rb_list_banks(state = "OR")` | `rb_get("banks", state = "OR")` |
| `rb_get_bank(17)` | `rb_get("banks", id = 17)` |
| `rb_get_banks(c(17, 100))` | `rb_get("banks", id = c(17, 100))` |
| `rb_list_ilf_programs()` | `rb_get("ilf")` |
| `rb_get_ilf_program(5)` | `rb_get("ilf", id = 5)` |
| `rb_epa_footprints()` | `rb_epa("footprints")` |
| `rb_epa_banks()` | `rb_epa("banks")` |
| `rb_download_potential_credits()` | `rb_download("potential_credits")` |

## RIBITS API Endpoints

Base URL: `https://ribits.ops.usace.army.mil/ords/RI/public/`

### List Endpoints

| Endpoint | Description | Parameters |
|----------|-------------|------------|
| `bank_site_list/` | List all banks | state, district, fieldoffice, noaaregion, kind_of_bank |
| `ilf_program_list/` | List ILF programs | state, district, fieldoffice, noaaregion |
| `umbrella_instrument_list/` | List umbrellas | state, district, fieldoffice, noaaregion |
| `wqt_project_list/` | List WQT projects | state, district |

### Data Endpoints

| Endpoint | Description | Parameters |
|----------|-------------|------------|
| `bank_site_data/` | Bank details | bank_id, show_ledger, show_footprint, show_service_area, show_contacts |
| `ilf_program_data/` | ILF details | program_id, show_ledger, show_service_area, show_contacts |
| `umbrella_instrument_data/` | Umbrella details | umbrella_instrument_id, show_ledger, show_service_area, show_contacts |

### Location Endpoints

| Endpoint | Description | Parameters |
|----------|-------------|------------|
| `getmitbanksbylocation_47/` | Banks near location | param0 (lon), param1 (lat) |
| `getprogramsbylocation_47/` | Programs near location | param0 (lon), param1 (lat) |

### Other Endpoints

| Endpoint | Description |
|----------|-------------|
| `find_credits_excel/{id}` | Download credits as Excel |

## EPA ArcGIS Layers

Base URL: `https://geopub.epa.gov/arcgis/rest/services/NEPAssist/RIBITS/MapServer`

| ID | Layer Name | rb_epa() Argument |
|----|------------|-------------------|
| 0 | USACE RIBITS Banks | `"banks"` |
| 2 | Approved Banks | `"banks"` (default) |
| 3 | Pending Banks | - |
| 4 | Terminated Banks | - |
| 5 | Bank Footprint | `"footprints"` |
| 6 | Bank Service Area | `"service_areas"` |
| 7 | USACE RIBITS ILFs | `"ilf_programs"` |
| 8 | ILF Program Service Areas | `"ilf_service_areas"` |
| 9 | ILF Approved | - |
| 10 | ILF Pending | - |
| 11 | ILF Terminated | - |
| 12 | USACE Districts | `"districts"` |

## Data Fields Reference

### Bank List Fields

Returned by `rb_get("banks")`:

| Field | Type | Description |
|-------|------|-------------|
| `bank_id` | integer | Unique bank identifier |
| `bank_name` | string | Bank name |
| `bank_status_name` | string | Status (Approved, Pending, Terminated) |
| `kind_of_bank_name` | string | Type (Standard, ILF, Umbrella, NRDA) |
| `state_abbrev_list` | string | States (comma-separated) |
| `district` | string | USACE district |
| `field_office` | string | FWS field office |
| `total_potential_credits` | numeric | Total potential credits |
| `total_available_credits` | numeric | Available credits |
| `longitude` | numeric | Bank longitude |
| `latitude` | numeric | Bank latitude |

### Bank Detail Fields

Returned by `rb_get("banks", id = X)`:

| Field | Type | Description |
|-------|------|-------------|
| `bank_id` | integer | Unique identifier |
| `bank_name` | string | Full bank name |
| `bank_status_name` | string | Current status |
| `site_name` | string | Site name |
| `approval_date` | date | Date approved |
| `establishment_date` | date | Date established |
| `comments` | list | Service area comments (nested) |
| `permittee_list` | list | Permittees (nested) |
| `ledger` | list | Transaction history (if requested) |
| `footprint` | GeoJSON | Footprint geometry (if requested) |
| `service_area` | GeoJSON | Service area geometry (if requested) |
| `contacts` | list | Contact information (if requested) |

### Ledger Transaction Fields

Returned by `rb_extract(bank, "ledger")`:

| Field | Type | Description |
|-------|------|-------------|
| `transaction_id` | integer | Unique transaction ID |
| `transaction_type` | string | Type: Init, Sell, Transfer, Adjust, etc. |
| `transaction_date` | date | Date of transaction |
| `credits` | numeric | Number of credits |
| `acres` | numeric | Acreage |
| `credit_classification` | string | Classification (Wetlands, Streams, Species) |
| `credit_type_list` | string | Credit type (Wetland, Stream, Species) |
| `jurisdiction` | string | Federal or State |
| `is_purchased` | string | Yes/No |
| `is_transferred` | string | Yes/No |
| `permittee_name` | string | Permittee name (if sold) |
| `permit_list` | string | Permit numbers (if sold) |
| `impact_huc_list` | string | Impact HUC codes |
| `comment` | string | Transaction comment |

### Transaction Types

| Type | Description |
|------|-------------|
| `Init` | Initial credit release |
| `Sell` | Credit sale to permittee |
| `Transfer` | Credit transfer between banks |
| `Adjust` | Credit adjustment |
| `Forfeit` | Credit forfeiture |
| `Withdraw` | Credit withdrawal |

### WQT Project Fields

| Field | Type | Description |
|-------|------|-------------|
| `wqt_project_id` | integer | Unique ID |
| `wqt_project_name` | string | Project name |
| `project_status_name` | string | Status |
| `project_cat_name` | string | Category |
| `state_abbrev_list` | string | States |
| `total_avail_phos` | numeric | Available phosphorus credits |
| `total_pend_of_phos` | numeric | Pending phosphorus credits |
| `total_potential_of_phos` | numeric | Potential phosphorus credits |

## Manual Download Reports

### Reports Available

| Report | Description | API Alternative |
|--------|-------------|-----------------|
| Credit Classification | Credits by type/jurisdiction | `rb_get_report_data()` |
| Potential Credits | Credits by mitigation method | **None - Manual only** |
| Credit Tracking | Transaction details | `rb_get(..., ledger = TRUE)` |
| Credit Withdrawal | Withdrawal details | Partial via ledger |
| Bank Summary | Bank overview | `rb_get("banks")` |
| Service Area Comments | Service area notes | `rb_get_report_data()` |

### Report URLs

| Report | Direct URL |
|--------|-----------|
| Credit Classification | `f?p=107:357:::::IR_71005` |
| Potential Credits | `f?p=107:357:::::IR_72659` |
| Credit Tracking | `f?p=107:357:::::IR_72644` |
| Credit Withdrawal | `f?p=107:12` |
| Bank Summary | `f?p=107:17` |

## Data Availability Matrix

| Data Category | API | Manual Download |
|---------------|-----|-----------------|
| Bank metadata | ✓ | |
| Bank locations | ✓ | |
| Ledger transactions | ✓ | |
| Credit classification | ✓ | ✓ |
| Permittee/permit info | ✓ | |
| Service area comments | ✓ | |
| Footprint geometry | ✓ | |
| Service area geometry | ✓ | |
| **Mitigation type breakdown** | ✗ | ✓ |
| Transfer tracking details | | ✓ |

## Filter Reference

### State Codes

Standard two-letter abbreviations: AL, AK, AZ, AR, CA, CO, CT, DE, FL, GA, HI, ID, IL, IN, IA, KS, KY, LA, ME, MD, MA, MI, MN, MS, MO, MT, NE, NV, NH, NJ, NM, NY, NC, ND, OH, OK, OR, PA, RI, SC, SD, TN, TX, UT, VT, VA, WA, WV, WI, WY

### USACE Districts

Examples: Portland, Sacramento, San Francisco, Los Angeles, Albuquerque, Galveston, Fort Worth, Tulsa, Little Rock, Memphis, Vicksburg, New Orleans, Mobile, Jacksonville, Savannah, Charleston, Wilmington, Norfolk, Baltimore, Philadelphia, New York, New England, Buffalo, Detroit, Chicago, St. Paul, Omaha, Kansas City, Rock Island

### Bank Types (kind)

- `Standard` - Standard mitigation bank
- `ILF` - In-Lieu Fee program site
- `Umbrella` - Umbrella bank
- `NRDA` - Natural Resource Damage Assessment bank

### Bank Status

- `Approved` - Active, approved bank
- `Pending` - Application pending
- `Terminated` - No longer active
- `Suspended` - Temporarily suspended
