---
title: "Getting Started with RIBITSr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with RIBITSr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

**RIBITSr** makes it easy to access data from RIBITS (Regulatory In-lieu fee and Bank Information Tracking System), the U.S. Army Corps of Engineers' national database for tracking mitigation banks and in-lieu fee programs.

### The RIBITSr Philosophy

RIBITSr follows a simple principle:

> **Get everything, then filter with tidyverse.**

Instead of complex download options, RIBITSr gives you the maximum amount of data by default (~85 transaction columns!), then lets you shape it using familiar tidyverse tools.

## Installation

```{r install}
# Install from GitHub
# install.packages("pak")
pak::pak("alexanderdhond/RIBITSr")

# Load the package
library(RIBITSr)
library(dplyr)  # For filtering and analysis
```

## The Simple Workflow

### Step 1: Get Everything

```{r simple-example}
library(RIBITSr)

# Get all data for California (one line!)
ca <- ribits(state = "CA")

# What did we get?
ca
#> <ribits_data: 128 banks>
#> ✔ 128 banks with summaries
#> ✔ 4,521 transactions (~85 columns, harmonized from 3 sources)
#> ✔ 342 contacts
#> ✔ 256 spatial geometries
```

That's it! One function call gets you:
- All bank attributes from 3 sources (API, EPA, CSV)
- Comprehensive transaction data (~85 columns)
- Spatial data (footprints, service areas)
- All automatically harmonized

### Step 2: Filter with Tidyverse

```{r filter-example}
library(dplyr)

# Find approved banks with high credit availability
high_credit_banks <- ca$banks %>%
  filter(
    bank_status == "Approved",
    available_credits > 100,
    total_acres > 50
  ) %>%
  select(bank_id, bank_name, bank_type, available_credits, total_acres) %>%
  arrange(desc(available_credits))

high_credit_banks
#> # A tibble: 23 × 5
#>    bank_id bank_name                  bank_type            available_credits total_acres
#>      <int> <chr>                      <chr>                            <dbl>       <dbl>
#>  1     123 Sacramento River Bank      Private Commercial               1245.        456
#>  2     456 Napa Valley Conservation   Non-Profit                       892.         234
#>  3     789 Bay Delta Mitigation       Public                           654.         678
#>  ...
```

## Understanding the Data Structure

The `ribits()` function returns a `ribits_data` object with four main components:

```{r data-structure}
# Main bank information (1 row per bank)
ca$banks
#> # A tibble: 128 × 67
#>    bank_id bank_name          bank_status district    state_list total_acres ...
#>      <int> <chr>              <chr>       <chr>       <chr>            <dbl>
#>  1      17 Vernal Pool Ranch  Approved    Sacramento  CA                 567
#>  2     100 Cosumnes River...  Approved    Sacramento  CA                 1234
#>  ...

# Transaction data (many rows per bank, ~85 columns!)
ca$transactions
#> # A tibble: 4,521 × 85
#>    bank_id transaction_id transaction_date permittee_name debits huc8 ...
#>      <int> <chr>          <date>           <chr>          <dbl> <chr>
#>  1      17 TXN-2023-001   2023-05-15       Acme Dev        12.5 18020111
#>  2      17 TXN-2023-045   2023-06-22       Smith Bros      8.3  18020111
#>  ...

# Spatial data (sf object ready for mapping)
ca$geometry
#> Simple feature collection with 256 features
#>    bank_id geometry_type      geometry
#>      <int> <chr>              <POLYGON>
#>  1      17 footprint          POLYGON((...)
#>  2      17 service_area       POLYGON((...)
#>  ...

# Metadata (sources, discrepancies, timing)
ca$.meta
#> $sources
#> [1] "api" "epa" "csv"
#>
#> $timing
#> $timing$duration_secs
#> [1] 45.3
#>
#> $discrepancies
#> # 3 minor discrepancies found and resolved
```

## Common Tasks

### Filter Banks by Attributes

```{r filter-banks}
# Banks in specific watershed
huc8_banks <- ca$banks %>%
  filter(grepl("18020111", huc8_list))

# Banks established after 2015
recent_banks <- ca$banks %>%
  filter(year_established > 2015)

# Banks by type
private_banks <- ca$banks %>%
  filter(bank_type == "Private Commercial")

# Combine multiple conditions
large_approved <- ca$banks %>%
  filter(
    bank_status == "Approved",
    total_acres > 200,
    available_credits > 50
  )
```

### Analyze Transactions

```{r analyze-transactions}
# Recent transaction activity
recent_activity <- ca$transactions %>%
  filter(transaction_date >= as.Date("2020-01-01")) %>%
  group_by(bank_id, bank_name) %>%
  summarize(
    n_transactions = n(),
    total_debits = sum(debits, na.rm = TRUE),
    avg_transaction = mean(debits, na.rm = TRUE),
    first_transaction = min(transaction_date),
    last_transaction = max(transaction_date)
  ) %>%
  arrange(desc(total_debits))

# Top permittees by credit usage
top_permittees <- ca$transactions %>%
  group_by(permittee_name) %>%
  summarize(
    total_credits = sum(debits, na.rm = TRUE),
    n_transactions = n(),
    n_banks_used = n_distinct(bank_id)
  ) %>%
  filter(total_credits > 0) %>%
  arrange(desc(total_credits)) %>%
  head(10)
```

### Summarize by District

```{r summarize-district}
# Credit availability by district
district_summary <- ca$banks %>%
  filter(bank_status == "Approved") %>%
  group_by(district) %>%
  summarize(
    n_banks = n(),
    total_available = sum(available_credits, na.rm = TRUE),
    total_released = sum(released_credits, na.rm = TRUE),
    avg_acres = mean(total_acres, na.rm = TRUE),
    median_acres = median(total_acres, na.rm = TRUE)
  ) %>%
  arrange(desc(total_available))
```

### Join Banks and Transactions

```{r join-example}
# Get bank details with transaction counts
bank_activity <- ca$transactions %>%
  group_by(bank_id) %>%
  summarize(
    n_transactions = n(),
    total_debits = sum(debits, na.rm = TRUE)
  ) %>%
  left_join(ca$banks, by = "bank_id") %>%
  select(bank_id, bank_name, district, n_transactions, total_debits, available_credits)
```

## Working with Spatial Data

All spatial data is returned as sf objects, ready for mapping:

```{r spatial-analysis}
library(sf)
library(ggplot2)

# Map bank footprints
ggplot(ca$geometry %>% filter(geometry_type == "footprint")) +
  geom_sf(aes(geometry = geometry), fill = "lightblue", alpha = 0.6) +
  theme_minimal() +
  labs(title = "California Mitigation Bank Footprints")

# Find banks near a specific location
library(sf)

# Convert banks to sf object
ca_banks_sf <- st_as_sf(ca$banks,
                        coords = c("longitude", "latitude"),
                        crs = 4326)

# Define target location (Sacramento)
sacramento <- st_point(c(-121.4944, 38.5816)) %>%
  st_sfc(crs = 4326)

# Find banks within 50km
nearby_banks <- ca_banks_sf %>%
  mutate(
    distance_km = as.numeric(st_distance(geometry, sacramento)) / 1000
  ) %>%
  filter(distance_km < 50) %>%
  arrange(distance_km)
```

## Options for Different Needs

By default, `ribits()` gives you **everything**. But you can customize:

### Speed vs Completeness

```{r speed-options}
# Maximum data (default) - comprehensive transactions (~85 columns)
full <- ribits(state = "OR")

# Faster - basic transactions (~20 columns)
basic <- ribits(state = "OR", transactions = "basic")

# Fastest - no transactions
fast <- ribits(state = "OR", transactions = "none", spatial = FALSE)
```

### Different Data Types

```{r data-types}
# ILF programs
ilf_programs <- ribits(type = "ilf", state = "CA")

# Umbrella instruments
umbrellas <- ribits(type = "umbrellas", state = "FL")

# Specific banks by ID
my_banks <- ribits(ids = c(17, 100, 345))
```

## Data Quality

Check data quality and see how sources were harmonized:

```{r quality-checks}
# Check for conflicts between sources
discrepancies(ca)
#> # A tibble: 3 × 5
#>   bank_id field          api_value  epa_value  resolution
#>     <int> <chr>          <chr>      <chr>      <chr>
#> 1      17 total_acres    567        565        Used API (more recent)
#> 2     100 bank_status    Approved   Pending    Used API (authoritative)

# See how conflicts were resolved
resolutions(ca)
#> All discrepancies resolved using priority: api > csv > epa

# Before downloading, check coverage
rb_check(state = "WA")
#> Data Coverage: state=WA
#> ✔ Banks in API: 45
#> ✔ Banks in EPA: 38 (84% of API)
#> ✔ With footprints: 35 (78%)

# After downloading, check quality
wa <- ribits(state = "WA")
rb_check(wa)
#> Data Quality Report
#> ✔ 45 banks
#> ✔ bank_name: 100% complete
#> ✔ total_acres: 96% complete
```

## Discovery Tools

Explore what data is available:

```{r discovery}
# See all available data sources
rb_info()
#> RIBITSr Data Catalog
#>
#> Quick Start:
#>   ca_data <- ribits(state = "CA")
#>
#> Data Sources:
#>   - RIBITS API: Real-time data
#>   - EPA ArcGIS: Spatial layers
#>   - CSV Reports: Official records
```

## Real-World Example: Watershed Analysis

Let's analyze mitigation banking activity in the Sacramento-San Joaquin Delta (HUC8: 18020104):

```{r watershed-example}
library(RIBITSr)
library(dplyr)
library(ggplot2)

# Get all California data
ca <- ribits(state = "CA")

# Filter to Delta watershed
delta_banks <- ca$banks %>%
  filter(grepl("18020104", huc8_list)) %>%
  select(bank_id, bank_name, bank_status, available_credits, total_acres)

# Get transactions for these banks
delta_transactions <- ca$transactions %>%
  filter(bank_id %in% delta_banks$bank_id) %>%
  mutate(year = lubridate::year(transaction_date))

# Analyze annual activity
annual_activity <- delta_transactions %>%
  filter(year >= 2015, year <= 2024) %>%
  group_by(year) %>%
  summarize(
    n_transactions = n(),
    total_credits = sum(debits, na.rm = TRUE),
    n_banks_active = n_distinct(bank_id)
  )

# Visualize trends
ggplot(annual_activity, aes(x = year, y = total_credits)) +
  geom_line(size = 1.2, color = "steelblue") +
  geom_point(size = 3) +
  labs(
    title = "Mitigation Banking Activity in Sacramento-San Joaquin Delta",
    subtitle = "HUC8: 18020104",
    x = "Year",
    y = "Total Credits Used"
  ) +
  theme_minimal()
```

## Configuration

Customize RIBITSr behavior:

```{r configuration}
# View current settings
rb_config()

# Enable persistent caching (survives R restart)
rb_config(
  use_persistent_cache = TRUE,
  cache_max_age_days = 7  # Refresh weekly
)

# For slow networks
rb_config(
  max_retries = 5,
  timeout = 120,
  rate_limit = 2  # Slower rate
)

# Reset to defaults
rb_config(reset = TRUE)
```

## Next Steps

Now that you've learned the basics:

1. **Explore your data** - Use dplyr and tidyr to filter and summarize
2. **Map your results** - Use sf and ggplot2 for spatial visualization
3. **Configure as needed** - Adjust caching and network settings with `rb_config()`
4. **Check data quality** - Use `rb_check()` and `discrepancies()` to understand your data

For more advanced usage, see:
- `vignette("workflow-examples")` - Common analysis patterns
- `vignette("configuration-guide")` - Detailed configuration options
- `?ribits` - Full function documentation

## Getting Help

- Function help: `?ribits`
- All vignettes: `browseVignettes("RIBITSr")`
- Report issues: [GitHub](https://github.com/alexanderdhond/RIBITSr/issues)
