---
title: "Getting Started with RIBITSr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with RIBITSr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

RIBITSr provides a streamlined interface to the RIBITS (Regulatory In-lieu fee 
and Bank Information Tracking System) API. RIBITS is the U.S. Army Corps of 
Engineers' national database for tracking mitigation banks and in-lieu fee programs.
This package allows you to:
- Query mitigation banks, ILF programs, and umbrella instruments
- Filter by state, district, status, and more
- Access transaction ledgers, footprints, and service areas
- Download spatial data from EPA's ArcGIS MapServer
- Handle manual report downloads for data not available via API

## Installation

```{r install}
# Install from GitHub
devtools::install_github("alexdhond/RIBITSr")

# Load the package
library(RIBITSr)
```

## Quick Start: The Unified API

> **Note:** RIBITSr now provides a streamlined `ribits()` function that automatically
> harmonizes data from multiple sources (API, EPA ArcGIS, CSV reports). For most use
> cases, `ribits()` is the recommended starting point. See the README for examples.
>
> This vignette documents the lower-level API functions (`rb_get()`, `rb_epa()`, etc.)
> which provide more granular control when needed.

RIBITSr provides the following main functions:

| Function | Purpose | When to Use |
|----------|---------|-------------|
| `ribits()` | **Recommended** - Auto-harmonized data from all sources | Most use cases, simplest API |
| `rb_get()` | Get banks, ILF programs, umbrellas, or WQT projects | Need API-only data or specific control |
| `rb_epa()` | Get spatial data from EPA ArcGIS | Need EPA spatial layers only |
| `rb_download_report()` | Download CSV reports directly (no browser needed) | Need specific CSV reports |
| `rb_read()` | Read downloaded CSV files | Working with downloaded data |

### Exploring Available Options

Each function shows help when called without arguments:

```{r help}
# See available data types and filters
rb_get()

# See available spatial layers
rb_epa()

# See available reports to download
rb_reports()

# Read any CSV file (auto-detects type)
data <- rb_read("path/to/file.csv")
```

## Querying Banks

### List Banks with Filters

```{r list-banks}
# Get all Oregon banks
or_banks <- rb_get("banks", state = "OR")
head(or_banks)

# Get banks in Sacramento district
sac_banks <- rb_get("banks", district = "Sacramento")

# Combine filters
ca_approved <- rb_get("banks", 
                       state = "CA", 
                       status = "Approved",
                       kind = "Standard")
```

### Available Filters

| Filter | Description | Example Values |
|--------|-------------|----------------|
| `state` | State code | "CA", "OR", "TX" |
| `district` | USACE district | "Portland", "Sacramento", "Galveston" |
| `field_office` | FWS field office | "Oregon Fish and Wildlife Office" |
| `noaa_region` | NOAA region | "West Coast" |
| `kind` | Bank type | "Standard", "ILF", "Umbrella", "NRDA" |
| `status` | Approval status | "Approved", "Pending", "Terminated" |

### Get Detailed Bank Data

When you need more than just the list, request specific data:

```{r get-bank}
# Get a single bank with ledger data
bank <- rb_get("banks", id = 17, ledger = TRUE)

# Get bank with all available data
bank_full <- rb_get("banks", id = 17, 
                    ledger = TRUE, 
                    footprint = TRUE,
                    service_area = TRUE,
                    contacts = TRUE)

# Get multiple banks at once
banks <- rb_get("banks", id = c(17, 100, 345), ledger = TRUE)
```

### Data Options

| Option | Description | Returns |
|--------|-------------|---------|
| `ledger = TRUE` | Transaction history | List of transactions |
| `footprint = TRUE` | Bank footprint geometry | GeoJSON/sf object |
| `service_area = TRUE` | Service area geometry | GeoJSON/sf object |
| `contacts = TRUE` | Contact information | List of contacts |

## Other Data Types

### ILF Programs

```{r ilf}
# List California ILF programs
ca_ilf <- rb_get("ilf", state = "CA")

# Get specific ILF program with details
ilf_program <- rb_get("ilf", id = 5, ledger = TRUE, service_area = TRUE)
```

### Umbrella Instruments

```{r umbrellas}
# List umbrella instruments in Portland district
umbrellas <- rb_get("umbrellas", district = "Portland")

# Get umbrella with full data
umbrella <- rb_get("umbrellas", id = 10, ledger = TRUE)
```

### Water Quality Trading Projects

```{r wqt}
# List WQT projects
wqt <- rb_get("wqt", state = "TX")
```

## Extracting Nested Data

Bank responses contain nested data. Use `rb_extract()` to pull them out:
```{r extract}
# Get bank with nested data
bank <- rb_get("banks", id = 17, ledger = TRUE, contacts = TRUE)

# Extract ledger as a tibble
ledger <- rb_extract(bank, "ledger")

# Extract contacts
contacts <- rb_extract(bank, "contacts")

# Extract specific contact types
sponsors <- rb_extract(bank, "sponsors")
pocs <- rb_extract(bank, "pocs")
```

### Bulk Extraction

```{r bulk}
# Bulk ledger extraction (from API)
ledgers <- rb_bulk_ledger(bank_ids = 1:50)

# Get transaction data (harmonized from multiple sources)
transactions <- rb_transactions(bank_ids = c(17, 100, 345))
```

## Spatial Data from EPA ArcGIS

EPA maintains a mirror of RIBITS spatial data on their ArcGIS MapServer.

```{r epa}
# Get bank footprints for Oregon
or_footprints <- rb_epa("footprints", state = "OR")

# Get service areas for specific banks
service_areas <- rb_epa("service_areas", bank_ids = c(17, 100))

# Get ILF program service areas
ilf_areas <- rb_epa("ilf_service_areas")

# Get USACE district boundaries
districts <- rb_epa("districts")
```

### Available Layers

| Layer | Description |
|-------|-------------|
| `banks` | Bank point locations (approved) |
| `footprints` | Bank footprint polygons |
| `service_areas` | Bank service area polygons |
| `ilf_programs` | ILF program point locations |
| `ilf_service_areas` | ILF program service areas |
| `districts` | USACE district boundaries |

## Location-Based Queries

Find banks or programs near a specific location:

```{r near}
# Find banks near Portland, OR
nearby_banks <- rb_near(lat = 45.5152, lon = -122.6784, type = "banks")

# Find ILF programs near the same location
nearby_programs <- rb_near(lat = 45.5152, lon = -122.6784, type = "programs")
```

## Manual Downloads (Programmatic)

Some detailed data is only available via CSV reports from the RIBITS website.
RIBITSr can now download these directly without a browser!

### Download Reports

```{r download}
# Download "Potential Credits" report directly to disk
file_path <- rb_download_report("potential_credits")

# Download to a specific directory
rb_download_report("credit_classification", download_dir = "data/")

# See all available report types
rb_reports()
```

### Read Downloaded Files

```{r read}
# Read the downloaded file (auto-detects type)
potential <- rb_read(file_path)

# Read any downloaded CSV report
credits <- rb_read("data/credit_classification.csv")
```

## Data Availability Summary

| Data | API | Manual Download |
|------|-----|-----------------|
| Bank metadata & list | ✓ | |
| Ledger transactions | ✓ | |
| Credit classification | ✓ | |
| Permittee/permit info | ✓ | |
| Service area comments | ✓ | |
| Footprint geometry | ✓ | |
| Service area geometry | ✓ | |
| **Mitigation type breakdown** | ✗ | ✓ |
| Transfer tracking details | | ✓ |

## Next Steps

- See `vignette("workflow-examples")` for complete analysis workflows
- See `vignette("api-reference")` for detailed API documentation
- See `data-raw/api_reference.md` for complete field reference
