---
title: "Workflow Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This vignette demonstrates complete workflows for common RIBITS analysis tasks.

## Workflow 1: State-Level Bank Analysis

Analyze all mitigation banks in a state, including credit availability and transactions.

```{r state-analysis}
library(RIBITSr)
library(dplyr)
library(ggplot2)

# Step 1: Get all banks in Oregon
or_banks <- rb_get("banks", state = "OR")

# Step 2: Summarize by status
or_banks %>%
  count(bank_status_name) %>%
  arrange(desc(n))

# Step 3: Get detailed data for approved banks
approved_ids <- or_banks %>%
  filter(bank_status_name == "Approved") %>%
  pull(bank_id)

# Step 4: Fetch ledgers for all approved banks
or_ledgers <- rb_bulk_ledger(bank_ids = approved_ids)

# Step 5: Analyze transaction patterns
or_ledgers %>%
  group_by(transaction_type) %>%
  summarize(
    n_transactions = n(),
    total_credits = sum(credits, na.rm = TRUE)
  )

# Step 6: Credit availability over time
or_ledgers %>%
  filter(transaction_type == "Sell") %>%
  mutate(year = lubridate::year(transaction_date)) %>%
  group_by(year) %>%
  summarize(credits_sold = sum(credits, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = credits_sold)) +
  geom_col() +
  labs(title = "Oregon Bank Credits Sold by Year")
```

## Workflow 2: Spatial Analysis with EPA Data

Combine RIBITS data with EPA spatial data for mapping.

```{r spatial-analysis}
library(RIBITSr)
library(sf)
library(ggplot2)

# Step 1: Get California bank locations
ca_banks <- rb_epa("banks", state = "CA")

# Step 2: Get footprints for those banks
ca_footprints <- rb_epa("footprints", state = "CA")

# Step 3: Get service areas
ca_service_areas <- rb_epa("service_areas", state = "CA")

# Step 4: Map banks by type
ggplot() +
  geom_sf(data = ca_service_areas, fill = "lightblue", alpha = 0.3) +
  geom_sf(data = ca_footprints, fill = "darkgreen", alpha = 0.5) +
  geom_sf(data = ca_banks, aes(color = KIND_OF_BANK), size = 2) +
  theme_minimal() +
  labs(title = "California Mitigation Banks",
       color = "Bank Type")

# Step 5: Calculate service area coverage
ca_service_areas %>%
  mutate(area_sq_km = as.numeric(st_area(geometry)) / 1e6) %>%
  summarize(total_coverage = sum(area_sq_km, na.rm = TRUE))
```

## Workflow 3: Credit Classification Analysis

Analyze credit types and classifications across banks.

```{r credit-classification}
library(RIBITSr)
library(dplyr)
library(tidyr)

# Step 1: Get banks in Portland district
pdx_banks <- rb_get("banks", district = "Portland")

# Step 2: Get credit classifications
classifications <- rb_credit_classifications(bank_ids = pdx_banks$bank_id)

# Step 3: Summarize by classification
classifications %>%
  group_by(credit_classification, jurisdiction) %>%
  summarize(
    n_banks = n_distinct(bank_id),
    total_credits = sum(credits, na.rm = TRUE)
  ) %>%
  arrange(desc(total_credits))

# Step 4: Pivot for comparison
classifications %>%
  group_by(bank_name, credit_classification) %>%
  summarize(credits = sum(credits, na.rm = TRUE)) %>%
  pivot_wider(names_from = credit_classification, values_from = credits)
```

## Workflow 4: Permittee and Permit Analysis

Track who is purchasing credits and for which permits.

```{r permittee-analysis}
library(RIBITSr)
library(dplyr)

# Step 1: Get banks with ledger data
bank_ids <- c(17, 100, 345, 500, 750)
banks <- rb_get("banks", id = bank_ids, ledger = TRUE)

# Step 2: Extract all ledgers
all_ledgers <- purrr::map_dfr(banks, ~ rb_extract(.x, "ledger"))

# Step 3: Filter to sales (credit purchases)
sales <- all_ledgers %>%
  filter(transaction_type == "Sell") %>%
  select(bank_id, transaction_date, credits, permittee_name, permit_list)

# Step 4: Top permittees by credits purchased
sales %>%
  group_by(permittee_name) %>%
  summarize(
    total_credits = sum(credits, na.rm = TRUE),
    n_transactions = n()
  ) %>%
  arrange(desc(total_credits)) %>%
  head(10)

# Step 5: Permits over time
sales %>%
  mutate(year = lubridate::year(transaction_date)) %>%
  count(year) %>%
  ggplot(aes(x = year, y = n)) +
  geom_line() +
  labs(title = "Credit Sales Transactions by Year")
```

## Workflow 5: ILF Program Comparison

Compare ILF programs across districts.

```{r ilf-comparison}
library(RIBITSr)
library(dplyr)

# Step 1: Get all ILF programs
all_ilf <- rb_get("ilf")

# Step 2: Compare by district
all_ilf %>%
  count(district, sort = TRUE)

# Step 3: Get details for specific programs
ca_ilf <- rb_get("ilf", state = "CA")
ca_ilf_details <- rb_get("ilf", id = ca_ilf$program_id[1:5], 
                          ledger = TRUE, service_area = TRUE)

# Step 4: Analyze ILF transactions
ilf_ledgers <- purrr::map_dfr(ca_ilf_details, ~ rb_extract(.x, "ledger"))

ilf_ledgers %>%
  group_by(transaction_type) %>%
  summarize(total_credits = sum(credits, na.rm = TRUE))
```

## Workflow 6: Potential Credits by Mitigation Type

This data requires manual download, but here's the complete workflow.

```{r mitigation-type}
library(RIBITSr)
library(dplyr)

# Step 1: Download the report (opens browser)
# b <- rb_download("potential_credits")
# In browser: Click Actions > Download > CSV
# b$close()

# Step 2: Read the downloaded file
potential <- rb_read("potential_credits")

# Step 3: Analyze mitigation methods
potential %>%
  group_by(mitigation_method) %>%
  summarize(
    n_banks = n_distinct(bank_id),
    total_acres = sum(acres, na.rm = TRUE),
    total_credits = sum(credits, na.rm = TRUE)
  ) %>%
  arrange(desc(total_credits))

# Step 4: Compare establishment vs preservation
potential %>%
  filter(mitigation_method %in% c("Establishment", "Preservation")) %>%
  group_by(mitigation_method, state) %>%
  summarize(credits = sum(credits, na.rm = TRUE)) %>%
  pivot_wider(names_from = mitigation_method, values_from = credits)
```

## Workflow 7: Finding Banks Near a Project Location

Find available mitigation credits near a specific project site.

```{r location-search}
library(RIBITSr)
library(dplyr)
library(sf)

# Step 1: Define project location (Portland, OR)
project_lat <- 45.5152
project_lon <- -122.6784

# Step 2: Find nearby banks
nearby <- rb_near(lat = project_lat, lon = project_lon, type = "banks")

# Step 3: Get service areas to check coverage
service_areas <- rb_epa("service_areas", bank_ids = nearby$bank_id)

# Step 4: Check which service areas contain the project
project_point <- st_sfc(st_point(c(project_lon, project_lat)), crs = 4326)

covering_areas <- service_areas %>%
  filter(st_contains(geometry, project_point, sparse = FALSE)[,1])

# Step 5: Get bank details for covering areas
if (nrow(covering_areas) > 0) {
  available_banks <- rb_get("banks", id = covering_areas$BANK_ID, ledger = TRUE)
  
  # Check credit availability
  purrr::map_dfr(available_banks, function(bank) {
    ledger <- rb_extract(bank, "ledger")
    tibble(
      bank_id = bank$bank_id,
      bank_name = bank$bank_name,
      total_potential = sum(ledger$credits[ledger$transaction_type == "Init"], na.rm = TRUE),
      total_sold = sum(ledger$credits[ledger$transaction_type == "Sell"], na.rm = TRUE),
      available = total_potential - total_sold
    )
  })
}
```

## Workflow 8: Complete Data Export

Export all data for a set of banks to files.

```{r export}
library(RIBITSr)
library(dplyr)
library(sf)
library(readr)

# Step 1: Define banks to export
bank_ids <- c(17, 100, 345, 500)

# Step 2: Get full data
banks <- rb_get("banks", id = bank_ids, 
                ledger = TRUE, 
                footprint = TRUE,
                service_area = TRUE,
                contacts = TRUE)

# Step 3: Extract and save ledgers
all_ledgers <- purrr::map_dfr(banks, ~ {
  ledger <- rb_extract(.x, "ledger")
  if (!is.null(ledger)) ledger$bank_id <- .x$bank_id
  ledger
})
write_csv(all_ledgers, "output/ledgers.csv")

# Step 4: Extract and save contacts
all_contacts <- purrr::map_dfr(banks, ~ {
  contacts <- rb_extract(.x, "contacts")
  if (!is.null(contacts)) contacts$bank_id <- .x$bank_id
  contacts
})
write_csv(all_contacts, "output/contacts.csv")

# Step 5: Get and save spatial data
footprints <- rb_epa("footprints", bank_ids = bank_ids)
st_write(footprints, "output/footprints.gpkg")

service_areas <- rb_epa("service_areas", bank_ids = bank_ids)
st_write(service_areas, "output/service_areas.gpkg")
```

## Workflow 9: Automated Data Harmonization

Use the built-in harmonization functions for consistent data.

```{r harmonize}
library(RIBITSr)

# Step 1: Set up download directory
rb_setup_download_dir("data/ribits_manual")

# Step 2: Check what manual downloads are needed
rb_download_status("data/ribits_manual")

# Step 3: Download any missing reports
# rb_download("potential_credits")
# rb_download("credit_classification")

# Step 4: Run auto-harmonization
harmonized <- rb_auto_harmonize(
  bank_ids = 1:100,
  manual_download_dir = "data/ribits_manual"
)

# Step 5: Validate the harmonized data
rb_validate_harmonized(harmonized)
```

## Tips and Best Practices

### Performance

```{r performance}
# Use bulk operations when possible
ledgers <- rb_bulk_ledger(bank_ids = 1:100)  # Better than loop

# Cache expensive queries
banks <- rb_get("banks", state = "CA")
saveRDS(banks, "cache/ca_banks.rds")
# Later: banks <- readRDS("cache/ca_banks.rds")
```

### Error Handling

```{r error-handling}
# Handle missing data gracefully
safe_get <- function(bank_id) {
  tryCatch({
    rb_get("banks", id = bank_id, ledger = TRUE)
  }, error = function(e) {
    message("Failed to get bank ", bank_id, ": ", e$message)
    NULL
  })
}

# Get multiple banks with error handling
banks <- purrr::map(bank_ids, safe_get) %>%
  purrr::compact()  # Remove NULLs
```

### Memory Management

```{r memory}
# Process large datasets in chunks
all_bank_ids <- rb_get("banks")$bank_id

chunk_size <- 100
results <- list()

for (i in seq(1, length(all_bank_ids), by = chunk_size)) {
  chunk_ids <- all_bank_ids[i:min(i + chunk_size - 1, length(all_bank_ids))]
  results[[length(results) + 1]] <- rb_bulk_ledger(bank_ids = chunk_ids)
  gc()  # Clean up memory
}

all_ledgers <- dplyr::bind_rows(results)
```
