---
title: "Configuration Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Configuration Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(RIBITSr)
```

# Introduction

RIBITSr provides extensive configuration options to control network behavior, caching, performance, and data quality. This guide covers all configuration options and provides practical examples for common use cases.

## Quick Overview

All configuration is managed through the `rb_config()` function:

```{r}
# View current configuration
rb_config()

# Change settings
rb_config(max_retries = 5, rate_limit = 3)

# Reset to defaults
rb_config(reset = TRUE)
```

# Network & Reliability Settings

## Retry Behavior

RIBITSr automatically retries failed requests with exponential backoff. Configure retry behavior for unreliable networks:

```{r}
# Increase retries for unreliable connections
rb_config(
  max_retries = 5,      # Default: 3
  retry_delay = 3       # Initial delay in seconds (default: 2)
)
```

**How it works:**
- Failed requests are automatically retried with exponential backoff
- Retry delays: 2s, 4s, 8s, 16s, 32s...
- Permanent errors (404, 401, 403) are not retried
- Network errors and server errors (5xx) are retried

**When to adjust:**
- Increase `max_retries` if working from an unreliable network
- Increase `retry_delay` if the RIBITS server is slow to respond
- Decrease both for faster failures when debugging

## Timeouts

Control how long to wait for responses:

```{r}
# Set longer timeout for large CSV downloads
rb_config(timeout = 300)  # 5 minutes (default: 60 seconds)
```

**Note:** CSV downloads automatically use a 300-second timeout regardless of this setting.

## Rate Limiting

Prevent overwhelming the RIBITS server:

```{r}
# More conservative rate limit
rb_config(rate_limit = 2)  # 2 requests/second (default: 5)

# More aggressive (use with caution)
rb_config(rate_limit = 10)  # 10 requests/second

# Disable rate limiting (not recommended)
rb_config(rate_limit = 0)
```

**Best practices:**
- Default (5 req/sec) works well for most use cases
- Reduce to 2-3 if you encounter 429 (rate limit) errors
- Never disable rate limiting unless testing locally

# Caching

## Session-Only Cache (Default)

By default, RIBITSr caches data only for the current R session:

```{r}
# First call downloads from RIBITS
banks <- ribits(state = "CA")

# Second call uses cache (fast!)
banks <- ribits(state = "CA")

# Cache cleared when R restarts
```

## Persistent Cache

Enable caching across R sessions for frequently-accessed data:

```{r}
# Enable persistent caching
rb_config(
  use_persistent_cache = TRUE,
  cache_max_age_days = 7     # Auto-delete files older than 7 days
)
```

**Cache locations:**
- Session cache: `tempdir()/ribits_cache/` (deleted on R exit)
- Persistent cache: User cache directory via `rappdirs::user_cache_dir("ribits", "RIBITSr")`

## Cache Management

```{r}
# Clear all cache
rb_clear_cache()

# Clear only CSV reports
rb_clear_cache("csv")

# Clear only name lookup table
rb_clear_cache("lookup")

# Silent clearing
rb_clear_cache(verbose = FALSE)
```

## Custom Cache Directory

Specify a custom cache location:

```{r}
# Set custom cache directory
Sys.setenv(RIBITS_CACHE_DIR = "/path/to/my/cache")

# Or in .Renviron:
# RIBITS_CACHE_DIR=/path/to/my/cache
```

# Data Quality Settings

## Source Priority

When harmonizing data from multiple sources, control which source is preferred for conflict resolution:

```{r}
# Prefer API over CSV over EPA (default)
rb_config(source_priority = c("api", "csv", "epa"))

# Prefer CSV over API (e.g., if CSV is known to be more accurate)
rb_config(source_priority = c("csv", "api", "epa"))
```

## Automatic Conflict Resolution

Control how discrepancies between sources are handled:

```{r}
# Automatically resolve conflicts using source priority (default)
rb_config(auto_resolve = TRUE)

# Preserve all conflicting values and let user decide
rb_config(auto_resolve = FALSE)
```

When `auto_resolve = FALSE`, the `rb_diagnose()` function will show all conflicts for manual inspection.

# Environment Variables

Configure RIBITSr via environment variables for reproducible workflows:

## Setting Environment Variables

Create or edit `.Renviron` in your home directory or project root:

```bash
# Network settings
RIBITS_MAX_RETRIES=5
RIBITS_RETRY_DELAY=3
RIBITS_TIMEOUT=120
RIBITS_RATE_LIMIT=3

# Caching
RIBITS_USE_PERSISTENT_CACHE=true
RIBITS_CACHE_DIR=/path/to/cache
RIBITS_CACHE_MAX_AGE_DAYS=14

# Data quality
RIBITS_AUTO_RESOLVE=true
```

Environment variables are loaded automatically when RIBITSr loads. They have lower priority than `rb_config()` calls.

## Checking Environment Variables

```{r}
# See what environment variables are set
Sys.getenv("RIBITS_MAX_RETRIES")
Sys.getenv("RIBITS_RATE_LIMIT")
```

# Common Use Cases

## Scenario 1: Unreliable Network Connection

You're working from a location with spotty internet:

```{r}
# Maximize reliability
rb_config(
  max_retries = 5,           # More retries
  retry_delay = 4,           # Longer delays
  timeout = 180,             # 3-minute timeout
  rate_limit = 2             # Slower requests
)
```

## Scenario 2: Daily Analysis Workflow

You run the same analysis daily and want to minimize download time:

```{r}
# Enable persistent caching
rb_config(
  use_persistent_cache = TRUE,
  cache_max_age_days = 1      # Refresh daily
)

# First run: downloads and caches
banks <- ribits(state = "CA", transactions = "comprehensive")

# Subsequent runs same day: uses cache (fast!)
banks <- ribits(state = "CA", transactions = "comprehensive")
```

## Scenario 3: Bulk Data Extraction

You need to download data for all 50 states:

```{r}
# Optimize for bulk operations
rb_config(
  max_retries = 3,
  rate_limit = 3,            # Conservative rate limit
  use_persistent_cache = TRUE,
  cache_max_age_days = 7
)

# Extract state by state with checkpointing
all_states <- c("AL", "AK", "AZ", ...)
results <- list()

for (state in all_states) {
  results[[state]] <- ribits(state = state, transactions = "comprehensive")

  # Save checkpoint
  saveRDS(results, "ribits_bulk_extraction.rds")
}
```

## Scenario 4: Production Environment

Running automated reports on a server:

```{r}
# Reliable, conservative settings
rb_config(
  max_retries = 5,
  retry_delay = 2,
  timeout = 300,
  rate_limit = 3,
  use_persistent_cache = TRUE,
  cache_max_age_days = 1,    # Refresh nightly
  auto_resolve = TRUE
)
```

Set via environment variables in production `.Renviron`:

```bash
RIBITS_MAX_RETRIES=5
RIBITS_RATE_LIMIT=3
RIBITS_USE_PERSISTENT_CACHE=true
RIBITS_CACHE_MAX_AGE_DAYS=1
```

## Scenario 5: Development & Testing

You're developing new analysis code and need fast iterations:

```{r}
# Speed over reliability
rb_config(
  max_retries = 1,           # Fail fast
  timeout = 30,              # Short timeout
  rate_limit = 10,           # Fast requests (local testing only!)
  use_persistent_cache = TRUE  # Avoid re-downloading
)
```

# Troubleshooting

## Issue: Requests Keep Timing Out

**Symptoms:**
- "Request timeout" errors
- Operations never complete

**Solutions:**
```{r}
# Increase timeout
rb_config(timeout = 300)

# Check network connection
check_ribits_connection(verbose = TRUE)
```

## Issue: Rate Limit Errors (429)

**Symptoms:**
- "Rate limit exceeded" errors
- 429 HTTP status codes

**Solutions:**
```{r}
# Reduce rate limit
rb_config(rate_limit = 2)

# Check current failures
rb_network_failures()
```

## Issue: Stale Cached Data

**Symptoms:**
- Data doesn't update when expected
- Missing recent transactions

**Solutions:**
```{r}
# Clear cache manually
rb_clear_cache()

# Or reduce cache age
rb_config(cache_max_age_days = 1)

# Or disable cache for this query
banks <- ribits(state = "CA", cache = FALSE)
```

## Issue: Name Matching Failures

**Symptoms:**
- Many banks have NA bank_id
- "Could not match name" warnings

**Solutions:**
```{r}
# Clear and rebuild name lookup cache
rb_clear_cache("lookup")

# Rebuild with fresh API data
lookup <- rb_build_name_lookup(include_csv = FALSE, cache = FALSE)

# Check for specific bank
rb_match_names(
  data.frame(name = "My Bank Name"),
  lookup,
  name_col = "name"
)
```

## Issue: Inconsistent Data Quality

**Symptoms:**
- Different values across sources
- Unexpected NAs or discrepancies

**Solutions:**
```{r}
# Diagnose the data
banks <- ribits(state = "CA", transactions = "comprehensive")
rb_diagnose(banks)

# Adjust source priority if needed
rb_config(source_priority = c("csv", "api", "epa"))

# Disable auto-resolution to inspect conflicts
rb_config(auto_resolve = FALSE)
```

# Monitoring & Diagnostics

## Check Network Failures

```{r}
# View all failed requests from current session
rb_network_failures()
#> Network failures (last 50):
#>   timestamp           | description                  | error
#>   2024-01-15 10:23:45 | CSV download: transactions  | HTTP 503
#>   2024-01-15 10:24:12 | API request: bank 123       | Timeout
```

## Test Connectivity

```{r}
# Test RIBITS API connection
check_ribits_connection(verbose = TRUE)
#> ✔ RIBITS API is reachable
#> ✔ Response time: 234ms

# Test specific bank
rb_get("banks", id = "SAJ-2009-00123")
```

## View Configuration

```{r}
# See all settings
rb_config()
#> RIBITSr Configuration:
#>   Network & Reliability:
#>     - max_retries: 3
#>     - retry_delay: 2 seconds
#>     - timeout: 60 seconds
#>     - rate_limit: 5 requests/second
#>   Caching:
#>     - use_persistent_cache: FALSE
#>     - cache_max_age_days: 30
#>   Data Quality:
#>     - source_priority: api, csv, epa
#>     - auto_resolve: TRUE
```

# Best Practices

## 1. Use Persistent Cache for Repeated Analysis

```{r}
# At start of analysis script
rb_config(use_persistent_cache = TRUE, cache_max_age_days = 7)
```

## 2. Conservative Rate Limiting in Production

```{r}
# Production environment
rb_config(rate_limit = 3)  # Be a good API citizen
```

## 3. Set Configuration Once at Package Load

In your project's startup script or `.Rprofile`:

```{r}
.First <- function() {
  if (requireNamespace("RIBITSr", quietly = TRUE)) {
    RIBITSr::rb_config(
      max_retries = 5,
      use_persistent_cache = TRUE,
      cache_max_age_days = 7
    )
  }
}
```

## 4. Use Environment Variables for Portability

Instead of hardcoding settings in scripts, use `.Renviron`:

```bash
# .Renviron
RIBITS_USE_PERSISTENT_CACHE=true
RIBITS_RATE_LIMIT=3
```

This makes your code portable across different environments.

## 5. Clear Cache Periodically

```{r}
# Weekly maintenance
rb_clear_cache()
```

Or rely on automatic cleanup:

```{r}
rb_config(cache_max_age_days = 7)
```

# Advanced Configuration

## Debugging Network Issues

Enable verbose output for debugging:

```{r}
# Enable verbose mode (internal setting)
.network_options$verbose <- TRUE

# Make requests
banks <- ribits(state = "CA")
#> → Downloading CSV: transactions_watershed
#> → Progress: 1.2 MB / 3.4 MB [35%]
#> ✔ Downloaded in 2.3s
```

## Custom Validation Thresholds

Transaction validation automatically checks for suspicious values. Validation results are stored in metadata:

```{r}
result <- ribits(state = "CA", transactions = "comprehensive")

# Check validation results
result$.meta$validation
#> $valid: TRUE
#> $warnings: character(0)
#> $stats:
#>   - credit_col: "credit"
#>   - credit_range: min=0.01, max=5432.10
#>   - source_breakdown: csv_watershed=1234, api=456, csv_ledger=789
```

## Error Classification

RIBITSr automatically classifies errors as retryable or permanent:

- **Retryable:** Network errors, timeouts, 5xx server errors, 429 rate limit
- **Permanent:** 404 not found, 401 unauthorized, 403 forbidden, other 4xx errors

This is handled automatically by the retry system.

# Summary

RIBITSr's configuration system provides:

- **Flexible network settings** for different connectivity scenarios
- **Smart caching** with session-only or persistent options
- **Data quality controls** for source prioritization
- **Environment variable support** for reproducible workflows
- **Automatic retry and error handling** for reliability

For most users, the defaults work well. Adjust settings based on your specific needs:

- **Unreliable network?** → Increase retries and timeout
- **Daily workflows?** → Enable persistent cache
- **Bulk extraction?** → Conservative rate limiting + persistent cache
- **Production?** → Environment variables + persistent cache

For help with specific issues, see the Troubleshooting section or file an issue on GitHub.
